### 2016-01-30 halo/text/abc.tex

\documentclass[12pt, preprint]{aastex}
\usepackage[breaklinks,colorlinks, urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{bm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\usepackage{algorithmic,algorithm}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{ colorlinks,
linkcolor=darkblue,
filecolor=darkgreen,
urlcolor=darkred,
citecolor=darkblue}

\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand{\todo}[1]{{\bf \textcolor{red}{ #1}}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\lang}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\vep}{\bm{\epsilon}}
\newcommand{\ep}{\epsilon}
\newcommand{\pars}{\vec{\theta}}
\newcommand{\dev}{\mathrm{d}}

\begin{document}
% We need a better definitely shorter title...
\title{Approximate Bayesian Computation in Large Scale Structure: Constraining the Galaxy-Halo Connection}
\begin{abstract}
\todo{This is not a paper about hod, let's not start with describing hod. I suggest that we talk about parameter estimation in large scale structure in general and then move on to hod. Constraining hod is a just the first step in testing our methodology.} The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies
populate dark matter halos that has had much success in reproducing large scale galaxy
clustering and understanding galaxy evolution within the context of the halo model.
Constraints on the parameters that describe the HOD model have so far been inferred using 
the standard practice of Bayesian inference with likelihood estimation that assumes a
functional Gaussian form of the likelihood.\todo{(Do we want to include analytic models for clustering statistics here? Yes, also we should say we use summary statistics that are not independent: clustering + group. Also there is no analytic model for the abundance of galaxy group richnesses)}  % I think we should keep it simple. There's no need to go into detail about everything in the abstract. This is already a pretty sizable abstract. 
Likelihood free inference methods such as the
Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever 
having to calculate the likelihood.

In this work, we present ABC in the context of constraining the parameters of the HOD model
using {\it Halotools}, a publicly available ... First, we set a mock galaxy catalog created
with HOD parameters from \cite{Zheng:2007aa} \todo{why do we site here?} as our data catalog. Then we use ABC on a set 
% To give it a false sense of legitimacy... 
of observables measured from forward modeled mock catalogs generated with {\it Halotools} 
in order to accurately infer constraints on the HOD parameters that 
reproduce the original parameters. \todo{Remark about the comparison between our ABC 
method constraints and the standard MCMC constraints.} Furthermore, since ABC does 
not require computing the likelihood, we include observables beyond galaxy clustering 
such as the galaxy number density and the galaxy group multiplicity function, for which
accurately constructing a likelihood is intractable. 
We demonstrate that adding these measurements to our set of observables leads to stronger
constraints on the HOD parameters, specifically the ones that govern the satellite galaxy
population. Moreover, the constraints we infer illustrate the feasibility of ABC in 
cosmological contexts.

\todo{Hogg writes here too}

\end{abstract}

\section{Introduction}

Cosmology was revolutionized in the 1990s with the introduction of likelihoods---%
pro\-ba\-bil\-ities for the data given the theoretical model and parameter settings---%
for combining data from different surveys and performing principled inferences of
the cosmological parameters (CITE CITE).
Nowhere has this been more true than in cosmic microwave background studies,
where it is nearly possible to analytically evaluate a likelihood function that
involves no (or minimal) approximations (CITE Wandelt and others).

Fundamentally, the tractability of likelihood functions in cosmology flows from
the fact that the initial conditions are exceedingly close to Gaussian in form,
and many sources of measurement noise are also Gaussian. (may as well cite evidence for this too)
Likelihood functions are easier to write down and evaluate when things are closer to Gaussian,
so they get easier as the length scales get larger and the structure formation gets
younger.
Hence the suitability of CMB data for likelihood analyses.

In large-scale structure studies, which have galaxies, quasars, or quasar absorption
systems as tracers, the likelihood cannot be Gaussian:
Even if the initial conditions are perfectly Gaussian, the growth of structure
creates non-linearities which are non-Gaussian, and the galaxies form
within the density field in some complex manner that is unknown (or modeled only effectively).
Even if the galaxies were a Poisson sampling of the density field (which they are demonstrably
not; CITE PEOPLE), it would be tremendously hard to write down even an approximate
likelihood function.

The standard approach is to make two strong assumptions, both demonstrably wrong:
The first is that the (density and) two-point function measurements constitute
sufficient statistics for the cosmological parameters.
That is, the assumption is that all of the information about the cosmological model is
encoded in the mean and variance statistics of the tracer population.
This first assumption is equivalent to assuming that the density field is a Gaussian process,
and is therefore closer to true on larger scales.

The second strong and wrong assumption is that the likelihood functon for the data can
be approximated by a pseudo-likelihood function that is a Gaussian probability density
in the space of the two-point correlation function estimate.
That is, the assumption is that a pseudo-likelihood of the data given by a squared distance
(scaled by an inverse covariance matrix) between an estimate of the correlation function
(the data) and a theoretical prediction of that same function will return sensible posterior
inferences.
This assumption cannot be correct (in detail) at any scale, since a correlation function,
being related to the variance of a continuous field,
must satisfy non-trivial positive-definiteness
requirements (it must take the form of a valid kernel function; CITE SOMETHING).
These requirements truncate function space such that the likelihood in that functon space
could never be Gaussian.
This second assumption becomes truer as the correlation function gets better measured, so
it is most true on intermediate scales (where neither shot noise nor cosmic variance blow up).
We expect it to be very strongly violated at the baryon acoustic scale in particular!

Neither of these two assumptions are required for cosmological inferences, however:
We have very sophisticated simulations and emulations of structure and galaxy formation.
Indeed, the theory of large-scale structure \emph{is} the output of these simulations;
we can simulate the data in great detail.
Therefore, we can simulate not just the one- and two-point statistics of the galaxies,
but also any higher order statistics that might be relevant.
Furthermore, we can see the distributions of these statistics, and where the data lie within
those distributions.
That is---in principle---we don't need to make these strong and wrong assumptions;
we can create---in principle---a likelihood function directly from (an enormous number of)
simulation outputs.

Of course, any naive approach to sufficiently simulating the data would be ruinously
expensive.
Fortunately, there are principled, (relatively) efficient methods for minimizing computation
and delivering correct posterior inferences,
using only a data simulator and some choices about statistics.
Here we use \todo{Introduce the names ABC and LFI with references}.

As we will see, the LFI approach---despite cleverness---will be very computationally expensive.
For context, it is worth remembering that even the traditional approach to large-scale structure,
outlined above, requires enormous numbers of structure- and galaxy-formation simulations as well,
to construct the covariance matrices that go into the pseudo-likelihood.
So although our costs will be high, there are no inexpensive methods to compare to.

\todo{MJ: What Hogg wrote above doesn't mesh with (that is, repeats differently) what you have written below, so can you go through and fix that?  Lots of what is below should remain, but some should get woven into what's above.}

% style note: Here and in what follows: --- separates parenthetical phrases, -- connects words that appear together in a relationship (like galaxy--galaxy) and (galaxy--halo), and - connects pairs of words into an adjective (like two-point).

Inference of the cosmological parameters and understanding the formation of galaxies within the context of large scale 
structure typically assumes the existence of dark matter halos (CITE) and therefore requires an understanding of the galaxy--halo connection (CITE). Halos are biased tracers of the large scale structure (CITE), the statistical properties of which are well understood from the study of dark matter simulations (12 CHANNELS PEOPLE, 12!), 
and the general assumption is that galaxies reside in dark matter Halos \todo{evidence for this claim or rather, evidence that supports this assumption}. Halo occupation distribution---HOD---is a probabilistic framework that describes how galaxies populate dark matter halos. The central tenet of halo occupation model is that the probability that a certain number of galaxies populate a dark matter halo is determined by the mass of the host halo. This statistical prescription for connection of galaxies and halos has been successful in reproducing the galaxy two-point clustering as well as the galaxy--matter cross correlation---galaxy--galaxy 
lensing---within the context of the halo model, and has been used in understanding the co-evolution of galaxies and halos (how? CITE) and putting constraints on the relation between the stellar mass and halo mass of galaxies and clusters (M-N ratio?) as well as cosmological parameters.

In the context of large scale clustering of galaxies, inference of the cosmological and halo model parameters are performed by treating some estimate of the galaxy--galaxy two-point correlation function as the data,
assuming an analytic model for the theoretical two-point correlation function prediction,
and a Gaussian likelihood connecting the two.
The accuracy of the analytic model of the galaxy clustering statistics hinges upon the accuracy of the fitting functions such as halo bias, halo mass function that are accurate to ten percent level. There are other simplifying assumptions encoded in these analytic models such as the scale independence of the halo bias. The weighted difference of the number density of the observed galaxies and the analytic number density is also added to this. In this work, we investigate the how accurately we can infer the model parameters by relaxing the underlying assumptions of the parameter inference methods.

Approximate bayesian computation (ABC) provides a \emph{rejected sampling} framework that alleviates the issues arising from approximating the likelihood of observing the large scale population of galaxies given the cosmological and halo model parameters. ABC approximates the posterior probability distribution function over the model parameters given the data---the angular positions of the galaxies on the sky and their redshifts---by drawing proposals from the prior over the model parameters, simulating the data with the proposals, and rejecting the proposals for which the distance between the data and the simulation (measured by a metric) is greater than a certain small threshold.

In practice, the rejected sampling method is replaced by a more efficient adaptive importance sampling operation called Population Monte Carlo (PMC). Initially, PMC rejects the draws from prior pdf with a relatively large threshold. In the consequent steps, importance sampling from the previous samples that passed the distance criteria is performed, and the threshold is updated adaptively. That is, after each iteration, the threshold is lowered according to the distribution of the distances.

Furthermore, This sampling formalism permits us to sample the posterior of the model by simulating the data via a forward generative model that accounts for all the complications and uncertainties associated with the survey data that are difficult to take into account in approximating the likelihood function. In the context of the large galaxy surveys, these complications may include masks, nontrivial survey geometry, and missing data as a result of fiber collision.

Rejected sampling in ABC requires a distance function that quantifies the closeness of the simulation (e.g. fiber collided galaxies) to the data (observed positions of the galaxies in the survey). In principle, this distance metric is a positive definite function  that compares various summary statistics of the data and those of the simulations. In this investigation, we focus on the two-point correlation, number density, and the group statistics (more specifically the abundance of the galaxy group richnesses).

Performing a full cosmological inference using this sampling method requires a simulator that, (I) takes a set of cosmological parameters as input and generates a snapshot of dark matter particles at a given redshift using an N-body simulation, (II) builds a catalog of dark matter halos for that snapshot of the N-body simulation using a halo finder algorithm, and (III) paints galaxies into host halos given a set of halo occupation model parameters.

Computationally, this process is extremely demanding. One way to make this problem more tractable is to separate it into two parts: \emph{emulation}, and \emph{simulation}. That is, one can use a set of N-body simulations with known cosmologies, and perform halo occupation simulations using the pre-built halo catalogs of those simulations. In order to compute the summary statistics corresponding to the samples from the parameter space for which there is no N-body simulation, one can interpolate between the summary statistics of the parameters for which we have available N-body simulations. This procedure is called emulation. Emulation has been introduced in the literature in coyote universe [CITE COYOTE].

In this work however, we narrow our focus on inferring the posterior probability over the parameters of the halo model using an already existing halo catalog of a dark matter particle N-body simulation with a fixed cosmology. That is we simplify the problem by leaving the problem of emulation for future investigations.

The calculations regarding populating the dark matter halos, computing the clustering statistics and finding groups in the galaxy mock catalogs are performed using ({\fontfamily{cmss}\selectfont Halotools}) which is an open-source package tailored for modeling galaxy-halo connection in large scale structure modeling ({\fontfamily{cmss}\selectfont http://halotools.readthedocs.org}).

This paper is structured as follows. In Section \ref{sec:method},
we discuss the ABC sampling method and the components of the Halo occupation modeling. We also give a brief overview of the summary statistics used in this investigation. In Section \ref{sec:halo} we give a brief overview of the forward modeling of the galaxy mock catalogs using halotools as well as generating a mock galaxy catalog as data. In Section \ref{sec:results} we explore the results of our parameter inference using different summary statistics. Finally, we discuss and conclude in Section \ref{sec:discussion}.

\section{Method}\label{sec:method}

%%%%%%%%%%%%%%%%%%%%% SECTION %%%%%%%%%%%%%%%%%%%%%%%%%%
% Approximate Bayesian Computation  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Approximate Bayesian Computation}

ABC is based on rejection sampling, which is a Monte Carlo method 
used to draw samples from a probability distribution that is difficult 
to directly sample. For inference, the probability distribution that 
we want to sample is the posterior probability distribution function.  
Before we present the theory and algorithm of ABC, we provide a brief
overview of rejection sampling.

Briefly, rejection sampling is a method that draws samples from a 
probability distribution $f(\alpha)$, which is difficult to 
directly sample, by sampling from the distribution $M g(\alpha)$ that 
satisfies $f(\alpha) < M g(\alpha)$. $g(\alpha)$ is an instrumental 
distribution where sampling is easier and $M > 1$.

More specifically, in the context of simulation-based inference, 
the main goal is to sample from the joint probability of a
simulation $X$ and parameters $\pars$ given observed data $D$, which 
from Bayes rule can be written as 
\beq
p(\pars, X | D) = \frac{P(D|X)p(X|\pars)\pi(\pars)}{\mathcal{Z}}
\eeq
where $\pi(\pars)$ is the prior probabilities over the parameters of 
interest and $\mathcal{Z}$ is the evidence, 
\beq
\mathcal{Z} = \int d\pars \; dX\; p(D|X) p(X|\pars) \pi(\pars). 
\eeq
Since $p(\pars, X | D)$ cannot be directly sampled, we use rejection 
sampling with instrumental distribution 
\beq
q(\pars, X) = p(X|\pars) \pi(\pars)
\eeq
and 
\beq
M = \frac{\mathrm{max}\; p(D|X)}{\mathcal{Z}} > 1.
\eeq
The choices of $q(\pars, X)$ and $M$ satisfy the condition 
\beq
p(\pars, X | D) < M q(\pars, X)
\eeq
so we can sample $p(\pars, X | D)$ by drawing ${\pars, X}$ from 
the $q(\pars, X)$ distribution. In practice, this is done 
by first drawing $\pars$ from the prior $\pi(\pars)$ then generating a 
simulation $X = f(\pars)$ via the forward model. Afterwards, ${\pars, X}$ 
is accepted if
\beq
u < \frac{p(\pars, X | D)}{M q(\pars, X)} = \frac{p(D|X)}{\mathrm{max}\;p(D|X)}
\eeq
where $u$ is drawn from $\mathtt{Uniform}[0,1]$.

At this point, ABC postulates that $p(D|X)$

The method requires a prior probability over the parameters of interest 
($\pi(\theta)$), a generative forward model of the data given a set of 
parameters ($f(\theta)$) and a distance metric $\rho(x,D)$ which measures 
the distance between the simulation $X = f(\theta)$ and the observed data $D$.

Technically in large scale structure studies, the observed data is the 
total angular and redshift positions of galaxies on the sky. However, 
in practice, this data is an intractably high dimensional object. As a result, we 
utilize statistical summaries of galaxy positions as our observed 'data' 
and the distance metric we use measures the closeness of the statistical 
summaries of observed galaxy positions versus the simulated galaxy 
positions. Common summary statistics used to encapsulate the galaxy position 
information include the number density of the galaxy sample, the two-point 
correlation function (2PCF), three-point correlation function, and galaxy 
group statistics.

Therefore the distance metric measures the closeness of statistical summaries of the data and those of the simulation. When the observed data-set under study is the positions of all galaxies observed in the survey, possible statistical summaries of the data are the number density of galaxies, two-point correlation function, three-point correlation function, galaxy group statistics (counts in cells), and so on.

Here, we provide a brief overview of rejection sampling algorithm and how it is used in a simulation-based inference. We want to draw samples from a probability distribution $p(\alpha)$. Now, let us assume that: (i) there exists an auxiliary density $q(\alpha)$ from which we can directly sample from, and (ii) there exist $M>1$ which is an appropriate bound on $p(\alpha)/q(\alpha)$. Given these assumptions, rejection sampling works in the following way. We draw a sample $\alpha$ from the auxiliary density $q(\alpha)$. Afterwards, we draw a random number $u$ from a uniform distribution between zero and one. Then we accept the draw $\alpha$ if the $u$ is less than the ratio $p(\alpha)/Mq(\alpha)$. This procedure is repeated until sufficient number of draws are accepted.

In the context of a simulation-based inference, we main goal is to 
sample from the joint probability of a simulation $X$ and parameters 
$\pars$ given the observed data $D$: 
\beq
p(\pars,X | D)
\eeq
According to Bayes rule $p(\pars, X | D)$ can be written as 
\beq
p(\pars,X | D)  = \frac{p(D|X)p(X|\pars)\pi(\pars)}{\mathcal{Z}}
\eeq
where $\mathcal{Z}$  is the evidence, 
\beq
\mathcal{Z} = \int \dev \pars \; \dev X \; p(D|X)p(X|\pars)\pi(\pars),
\eeq

There exists a constant $M>1$ 
\beq 
M = \frac{\mathrm{max}_{X} \; p(D|X)}{Z},
\eeq 
and auxiliary density $q(\pars ,X)$
\beq 
q(\pars ,X)=p(X|\pars )\pi(\pars ),
\eeq 
such that:
\beq 
p(\pars ,X|D)<Mq(\pars ,X).
\eeq 
Now in a simulation-based inference, we first draw a pair $\{\pars , X \}$ from the density $q(\pars , X)$. This requires drawing a sample $\pars$ from the prior $\pi(\pars)$, and generating a simulation $X=f(\pars)$ via the forward model. Then, we accept the pair $\{\pars , X \}$, if the following condition is met

\beq 
u \sim \mathrm{Uniform}[0,1] < \frac{p(\pars,X|D)}{Mq(\pars,X)}=\frac{p(D|X)}{\mathrm{max}_{X}p(D|X)}.
\eeq

ABC postulates that the probability of observing data $D$ given simulation $X$, $p(D|X)$ is controlled by a threshold $\epsilon$, and that this probability is one if the distance between the data and the simulation, $\rho(D,X)$ is less than $\epsilon$. That is

\beq 
p(D|X) = p_{\epsilon}(D|X) \equiv \begin{cases}
                                    1 \ \ \  \mathrm{if} \ \rho(D,X)<\epsilon \; , \\
                                    0 \ \ \ \mathrm{otherwise} \; .
                                    \end{cases}
\eeq

Thus the set of parameters $\pars$ is accepted if $\rho(D,f(\pars))<\epsilon$. In principle, the distance function measures the distance 
between the statistical summaries of the data and those of the simulation. 
The distance, in general, can be a vector, with each component representing the distance between one summary statistics of the data (e.g. variance) and the same summary statistics of the simulation. That is
\beq 
\rho(D,X) = \Big[\rho_{1}(S_{1}(D) , S_{1}(X)), ... , \rho_{M}(S_{M}(D) , S_{M}(X)) \Big],
\eeq 
where $M$ is the number of statistical summaries used in the inference, $S_{i}$ is the i-th summary statistics, and $\rho_{i}$ is a scalar function that measures the distance between the i-th summary statistics of the data and that of the simulation. In this case, the threshold $\epsilon$ is an $M$ dimensional vector.

We draw a large number of proposals from the prior distribution over the parameter space, 
and reject the samples for which the distance between the data and the simulation corresponding to that 
sample is more than a certain threshold $\bm{\epsilon}$. This requirement is satisfied \emph{if and only if}
every component of the distance vector is less than its corresponding threshold. Note that, each summary statistics 
has a separate threshold. This procedure is done iteratively until, we are left with a certain number
of samples that satisfy the distance criteria. The final ensemble approximates the posterior probability distribution over the parameters of the generative model of the data.

For a sufficiently small threshold, this algorithm samples from the posterior probability, $p(\pars | D)$, and for an arbitrarily large threshold, the algorithm samples from the prior $\pi(\pars)$. This rejection sampling technique requires repeated sampling from the prior, and a very small threshold to succeed.

In this investigation instead, we use Population Monte Carlo Algorithm (PMC) as our sampling technique. PMC is an iterative method that performs rejected sampling from a sequence of distributions $\{p_{1}(\pars),...,p_{T}(\pars)\}$.These distributions are paired with a sequence of decreasing distance thresholds $\{\epsilon_{1} , ... , \epsilon_{T}\}$.

At each iteration, $N$ proposals are drawn from a distribution over parameters, and a weight is assigned to each one. In the first step, the threshold is set to an arbitrarily large value. Therefore the algorithm makes $N$ draws from the prior and assigns equal weights to all the proposals.

In the consequent iterations, the distance threshold is set to a smaller value. A proposal is weighted-sampled from the previous sample. It is perturbed by a kernel which is set to the covariance of the parameters in the previous sample. A simulation is then generated by feeding the perturbed parameter into the forward model. The perturbed parameter is then accepted if the distance between the data and the generated simulation is less than the value of the threshold in that iteration. This procedure is repeated until $N$ particles are sampled. Then the weights are updated according to the kernel, the prior, and the previous set of weights.

\begin{algorithm}
\caption{The procedure for ABC-PMC}
\begin{algorithmic}[1]
%\STATE \DATA: D
%\STATE \RESULT: ABC posterior sample of $\pars$
\IF{$t=1:$}
%\STATE $\epsilon_t \gets \infty$
\FOR{$i=1,...,N$}
   \STATE // \emph{This loop can now be done in parallel for all i}
   \WHILE{$\rho(X,D)>\epsilon_t$}
   \STATE $\pars^{*}_{t} \gets \pi(\pars)$
   \STATE $X = f(\pars^{*}_{t})$
   \ENDWHILE
   \STATE $\pars^{(i)}_{t} \gets \pars^{*}_{t}$
   \STATE $w^{(i)}_{t} \gets 1/N$
\ENDFOR
\ENDIF
\IF{$t=2,...,T:$}
\FOR{$i=1,...,N$}
   \STATE // \emph{This loop can now be done in parallel for all i}
   \WHILE{$\rho(X,D)>\epsilon_t$}
   \STATE Draw $\pars^{*}_{t}$ from $\{\pars^{j}_{t-1}\}$ with probabilities $\{w^{j}_{t-1}\}$
   \STATE $\pars^{*}_{t} \gets K(\pars^{*}_{t},.)$
   \STATE $X = f(\pars^{*}_{t})$
   \ENDWHILE
   \STATE $\pars^{(i)}_{t} \gets \pars^{*}_{t}$
   \STATE $w^{(i)}_{t} \gets \frac{\pi(\pars^{(i)}_{t})}{\sum_{j=1}^{N}w_{t-1}^{(i)}K(\pars^{(j)}_{t-1},\pars^{(i)}_{t})}$
\ENDFOR
\ENDIF
\end{algorithmic}
\end{algorithm}

%\begin{algorithm}[H]
% \Data{this text}
% \Result{how to write algorithm with \LaTeX2e }
% initialization\;
% \While{not at end of this document}{
%  read current\;
%  \eIf{understand}{
%   go to next section\;
%   current section becomes this one\;
%   }{
%   go back to the beginning of current section\;
%  }
% }
% \caption{How to write algorithms}
%\end{algorithm}

\subsection{Forward model of the large scale galaxy populations}\label{sec:halo}
\subsubsection{halo occupation modeling}

\newcommand{\lcdm}{\Lambda {\rm CDM}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\mean}[2]{\left\langle#1 \vert {#2}\right\rangle}

%%% HALO OCCUPATION DISTRIBUTION %%%
\newcommand{\ngal}{N_{\mathrm{g}}}
\newcommand{\nsat}{N_\mathrm{s}}
\newcommand{\ncen}{N_\mathrm{c}}
\newcommand{\pnm}[2]{P(#1|#2)}

\newcommand{\mhalo}{M_{\rm h}}
\newcommand{\mvir}{M_\mathrm{vir}}

\newcommand{\dndmvir}{\frac{\dd n}{\dd\mvir}}
\newcommand{\dndmhalo}{\frac{\dd n}{\dd\mhalo}}
\newcommand{\dndmvirprime}{\frac{\dd n}{\dd\mvir'}}

%%% CLUSTERING STATISTICS %%%
\newcommand{\xigg}{\xi_{\mathrm{gg}}}
\newcommand{\xihh}{\xi_{\mathrm{hh}}}
\newcommand{\xiggr}{\xi_{\mathrm{gg}}(r)}
\newcommand{\xiggroneh}{\xi^{1h}_{\mathrm{gg}}(r)}
\newcommand{\xiggrtwoh}{\xi^{2h}_{\mathrm{gg}}(r)}
\newcommand{\ngalaxy}{\bar{n}_{\mathrm{g}}}

One of the most well-studied statistics of the galaxy distribution 
is $\xigg,$ the galaxy-galaxy correlation function. Mathematically, $\xigg$ encodes the
excess probability that a pair of galaxies in a sample are 
separated by three-dimensional distance $r.$ 
Galaxy clustering is straightforward to measure observationally \todo{(redshift space distortions?)}, 
and it is a highly sensitive probe of both the physics of galaxy evolution 
as well as the fundamental parameters of $\lcdm$ cosmology.

The assumption that galaxies reside in dark matter halos is the bedrock underlying 
all contemporary theoretical predictions for galaxy clustering. Since the 
clustering of dark matter halos can be predicted to high precision with N-body simulations, 
then knowledge of how galaxies populate and are distributed within halos 
is sufficient to predict $\xigg.$ Under this assumption, $\xigg$ can be formally decomposed
into a ``one-halo term" and a ``two-halo term'', 
referring to galaxy pairs that reside in the same halo vs. distinct halos, respectively. 
The two-halo term of $\xigg$ is a pair-number weighted average of $\xihh,$ 
the halo-halo correlation function, whereas the one-halo term is the pair-number 
weighted average of the intra-halo distribution of galaxies. 
Models for the galaxy-halo connection aiming to predict $\xigg$ must therefore 
provide theoretical means by which to calculate these weights and spatial distributions.

The Halo Occupation Distribution (HOD) is one of the most widely used 
 approaches to characterizing the galaxy-halo connection. 
The central quantity In the HOD is $\pnm{\ngal}{\mhalo},$ the probability that a halo of mass
$\mhalo$ hosts $\ngal$ galaxies. The first two moments of $\pnm$ contain the 
necessary information to calculate the one- and two-halo terms of the galaxy correlation function:

\begin{eqnarray}
\label{eq:onehaloterm}
1+\xiggroneh \simeq \frac{1}{4\pi{}r^{2}\ngalaxy^{2}}\int\dd\mhalo\dndmhalo\Xi_{\rm gg}(r|\mhalo) \times \mean{\ngal(\ngal-1)}{\mhalo},
\end{eqnarray} and

\begin{eqnarray}
\label{eq:twohaloterm}
\xiggrtwoh \simeq \xi_{\mathrm{mm}}(r)\left[\frac{1}{\ngalaxy}\int\dd\mhalo\dndmhalo \mean{\ngal}{\mhalo}b_{\mathrm{h}}(\mhalo)\right]^{2}
\end{eqnarray}

In Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}), $\ngalaxy$
is the galaxy number density,
$\dd\mathrm{n}/\dd\mhalo$ is the halo mass function, 
the spatial bias of dark matter halos is $b_{\mathrm{h}}(\mhalo),$ and
$\xi_{\rm mm}$ is the correlation function of dark matter.  
If we represent the spherically symmetric intra-halo distribution of galaxies 
by a unit-normalized $n_{\rm g}(r),$
then the quantity $\Xi_{\rm gg}(r)$ appearing in the above two equations 
is the convolution of $n_{\rm g}(r)$ with itself.

Employing Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) to 
calculate $\xigg(r)$ is the most common technical method used in HOD 
studies of galaxy clustering. 
However, this technique requires appeal to many simplifying assumptions 
in addition to the supposition that galaxies reside in halos. For example, 
Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) presume that 
the intra-halo galaxy distribution is spherically symmetric, and 
that the halo bias $b_{\mathrm{h}}(\mhalo)$ has no radial dependence. 
Moreover, these equations does not take halo exclusion into account. 
We refer the interested reader to \citet{cooray02,mo_vdb_white10,vdBosch13}, 
and references therein, for further information.

In addition to the above, in the vast majority of HOD studies the actual calculation of 
Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) relies on further assumption still. 
The quantities $b_{\mathrm{h}}(\mhalo),$ $\xi_{\mathrm{mm}}(r)$ and $\dndmhalo$ 
are commonly replaced by analytical fitting functions that have been calibrated 
against a suite of N-body simulations \citep[e.g.,][]{sheth_tormen01,smith03,tinker05,tinker10}.   
And yet, it is well established that the precision of these fitting functions is 
limited to the $\sim10\%$ level \citep[e.g.][]{tinker08}.

\todo{APH: Ok, so the above level of detail may be overkill for this paper. I figured that part of the argument we'd make is this: it is now time to start taking the precision of our predictions seriously, the ten-percent game is over. ABC is a big contribution to going beyond ten percent, and now that Halotools makes it possible to circumvent the need for these fitting functions, this opens up new limiting factors such as the hard Gaussianity assumption. So I see the above argument as part of the argument for using ABC. However, I'm not married to any of the above text, so just let me know if you want me to scrap some of it. Otherwise, I will next pick up from here and start talking about how Halotools relaxes the reliance on these approximations, and I will make explicit the point about how this new level of precision motivates the exploration of ABC.}

\todo{MJ: I think all of the text written above is necessary for making a case for simulation-based inference in large-scale structure modeling. So I'm in favor of keeping it for now.}

Dark matter halos are biased tracers of the underlying matter distribution of the universe. The halo model postulates that galaxies reside in dark matter halos. Halo Occupation distribution---HOD---is a powerful tool for description of the galaxy--halo connection. HOD provides a probabilistic prescription for how galaxies populate halos. According to (conventional) HOD modeling, the number of galaxies that populate a dark matter halo is dictated by the host halo mass $M_h$. The probability of a dark matter halo of mass $M_h$ being populated by $N_g$ galaxies---$P(N_g|M_h)$--- is a Poisson distribution, and is separable to two parts 
\beq
P(N_g|M_h) = P(N_\mathrm{cen}|M_h) + P(N_\mathrm{sat}|M_h),
\eeq
where $N_\mathrm{cen}$ is the number of central galaxies, and $N_\mathrm{sat}$ is the number of satellites. Given a halo catalog generated from the output of a cosmological N-body simulation, these probability density functions can be used to make draws for each halo to find its occupation number of central and satellite galaxies.

\subsubsection{Generating mock galaxy populations with halotools}

\subsubsection{summary statistics of galaxy populations}

\subsubsection{distance metrics}

\section{Experiments}\label{sec:data}

\subsection{Mock observations}

\newcommand{\gmf}{\mathcal{\zeta}_{\rm g}}

As a testbed of our method, we create a mock population of galaxies in the following way. We set the upper limit of the luminosity of the mock galaxies to -20. Then mock observations---positions of galaxies--- are generated by populating the pre-built halo catalogs in the bolshoi simulation according to \cite{Zheng:2007aa} HOD model, and the best-fit HOD parameters found by \citet{Zheng:2007aa} for SDSS galaxy samples with the same luminosity threshold.

Afterwards, summary statistics of the mock observations are computed. These summary statistics include the man number density $\ngalaxy$ the galaxy-galaxy correlation function $\xigg$, the group multiplicity function $\gmf$. These quantities are measured in the following way.

Co-moving number density of the mock galaxy catalog $\ngalaxy$ is calculated. Th radial bins at which $\xigg$ is computed are selected to match the radial bins used in [CITE ZEHAVI ETAL] clustering analysis of the SDSS DR7 galaxy sample. The measured $\xigg$ of the mock galaxy catalog is shown in Figure ~\ref{1}. The error-bars are measured from the diagonal elements of the error covariance matrix built from 500 mocks. That is, with the same set of true HOD parameters, 500 mock galaxy catalogs are generated and after computing the $\xigg$ for every mock catalog, the covariance matrix shown in Figure \ref{2pcf_cov} is calculated in the following way.

\begin{eqnarray} 
\mathrm{C}_{\xi}[i,j] &=& \frac{1}{N_{\mathrm{mocks}}-1}\sum_{k=1}^{N_{\mathrm{mocks}}} \Big[\xigg^{(k)}(r_{i})-\bar{\xi}_{\rm gg}(r_{i})\Big]\Big[\xigg^{(k)}(r_{j})-\bar{\xi}_{\rm gg}(r_{j}) \Big], \\
\bar{\xi}_{\rm gg}(r_{i}) &=& \frac{1}{N_{\mathrm{mocks}}}\sum_{k=1}^{N_{\mathrm{mocks}}} \xigg^{(k)}(r_{i}).
\end{eqnarray}

FoF group IDs of galaxies are derived by setting the projected linking length $b_{\perp}$ to 0.2 and the line-of-sight linking length $b_{\parallel}$ to 0.1. Then the galaxies with identical group IDs are grouped together. The group IDs are then binned into galaxy group richness bins. The group richness bins are selected in away that they match the ones used in [CITE Berlind etal] analysis of SDSS galaxy group identification. The resulting product is the histogram of the galaxy group richnesses. Group multiplicity function $\gmf$ of the mock observations is computed by divinding the number of galaxies in each richness bin by the volume of the simulation box 250 $\mathrm{Mpc}^3\mathrm{h}^{-3}$.

The measured $\gmf$ of the mock galaxy catalog is shown in figure ~\ref{2}. The covariance between the group multiplicity functions between two richness bins has two components. The first one is the sample covariance between the richness bins. This is computed from the group multiplicity function evaluations of a set of 500 galaxy mock catalogs. This covariance is computed in the following way:

\begin{eqnarray} 
\mathrm{C}_{\zeta}^{\mathrm{sample}}[i,j] &=& \frac{1}{N_{\mathrm{mocks}}-1}\sum_{k=1}^{N_{\mathrm{mocks}}} \Big[\gmf^{(k)}(N_{i})-\bar{\zeta}_{\rm g}(N_{i})\Big]\Big[\gmf^{(k)}(N_{j})-\bar{\zeta}_{\rm g}(N_{j}) \Big], \\
\bar{\zeta}_{\rm g}(N_{i}) &=& \frac{1}{N_{\mathrm{mocks}}}\sum_{k=1}^{N_{\mathrm{mocks}}} \gmf^{(k)}(N_{i}).
\end{eqnarray}

The second contribution to the error covariance matrix of $\gmf$ measurements stems from Poisson noise which can be approximated by the number of galaxies within each richness bin devided by the volume of the box. That is,

\begin{eqnarray} 
\mathrm{C}_{\zeta}[i,j] &=& \mathrm{C}_{\zeta}^{\mathrm{sample}}[i,j] + \mathrm{C}_{\zeta}^{\mathrm{poisson}}[i,j], \\
\mathrm{C}_{\zeta}^{\mathrm{poisson}}[i,j] &=& \delta_{i,j}\frac{\langle N_{i}\rangle^{2}}{V_{\mathrm{box}}},
\end{eqnarray}
where $V_{\mathrm{box}} = 250 \mathrm{Mpc}^3\mathrm{h}^{-3}$, and $\langle N_{i}\rangle$ is the estimated number of galaxies in the richness bin $i$.

\subsection{Tests}

%Plots showing 2pcf, gmf, ... of the mock data.
%How we compute the covariance matrix from mocks...
%Talk about error bars. Talk about poisson errors in group %populations .
%--------------------------------------------------------
% Covariance matrix of xi  
%--------------------------------------------------------
\begin{figure}
\includegraphics[width=0.5\textwidth]{xi_covariance_Mr20_Nmock500.png}
\caption{The reduce covariance matrix $C_{i,j}/\sqrt{C_{i,i} C_{j,j}}$ of 
the two-point correlation function, $\xi(r)$, estimated from 500 Halotools 
realizations. We describe the calculation of $\xi(r)$ in Section \todo{reference section}. \todo{binning and range of $r$}.}
\label{fig:2pcf_cov}
\end{figure}
%--------------------------------------------------------
% Covariance matrix of GMF 
%--------------------------------------------------------
\begin{figure}
  \includegraphics[width=0.5\textwidth]{gmf_covariance_Mr20_Nmock500.png}
\caption{The reduce covariance matrix $C_{i,j}/\sqrt{C_{i,i} C_{j,j}}$ of 
the Galaxy Multiplicity Function (GMF) estimated from 500 Halotools 
realizations. \todo{reference to GMF calculation in text}. 
\todo{binning and range of $r$}.}
\label{fig:gmf_cov}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%% SECTION: %%%%%%%%%%%%%%%%%%%%%%%
% SECTION: RESULTS 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec:results}
Show the posteriors...
Show evolution of the posterior errorbars with iteration. Show epsilon. Convice that we are converging ...
Draw samples from the posterior, populate the mock catalogs with those samples, and compute the summary statistcs with them 
and demonstrate that they match the data...

\subsection{Comparison to MCMC}
\todo{Apples to apples comparison between ABC and MCMC.}
ABC-PMC using nbar and xi vs MCMC
%--------------------------------------------------------
% ABC-PMC Posterior likelihoods using nbar and xi
%--------------------------------------------------------
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{nbar_xi_Mr20_Niter10000_Nburn100_mcmc_samples_test.png}
\caption{Posterior likelihood of HOD parameters $\log M_0$, 
$\sigma_{\log M}$, $\log M_{min}$, $\alpha$, and $\log M_1$ using the
standard Markov Chain Monte Carlo method as described in Section \todo{CITECITE}. 
\todo{THIS IS A SPACE MARKER NOT THE FINAL FIGURE.}}
\label{fig:post_abc_nbarxi}
\end{center}
\end{figure*}
%--------------------------------------------------------
% MC-MC Posterior likelihoods using nbar and xi
%--------------------------------------------------------
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{nbar_xi_Mr20_now_t11.png}
\caption{Posterior likelihood of HOD parameters $\log M_0$, 
$\sigma_{\log M}$, $\log M_{min}$, $\alpha$, and $\log M_1$ using ABC-PMC 
with distance criteria defined by the observables $\bar{n}$ and $\xi(r)$.
\todo{This is the 12th iteration. NOT THE FINAL.}}
\label{fig:post_abc_nbarxi}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%% SECTION: %%%%%%%%%%%%%%%%%%%%%%%
% ABC INCLUDING GMF
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{ABC Including GMF}
ABC-PMC results using observables beyond galaxy clustering and nbar(z) GMF
%--------------------------------------------------------
% ABC-PMC Posterior likelihoods using nbar, xi and GMF 
%--------------------------------------------------------
\todo{Insert ABC-PMC Posterior Likelihoods using $\bar{n}$, $\xi(r)$ and GMF 
once it finishes running}

\section{Discussion}\label{sec:discussion}
\todo{talk about cosmological inference and why this is good first step for likelihood free inference in large scale ...}

talk about halo model ...

talk about HOD and why it is too simple to be the right model ...

talk about assembly bias perhaps ... and how we can use the same machinery for detection of assemblt bias ...

talk about writing down a likelihood for gmf, tpcf, ... , and the Gaussianity assumption ....

talk about are results ...
talk about future directions ...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC-PMC Posterior likelihoods using nbar and GMF 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{nbar_gmf5_Mr20_now_t18.png}
\caption{Posterior likelihood of HOD parameters $\log M_0$, 
$\sigma_{\log M}$, $\log M_{min}$, $\alpha$, and $\log M_1$ using ABC-PMC 
with distance criteria defined by the observables $\bar{n}$ and GMF.}
\label{fig:post_abc_nbargmf}
\end{center}
\end{figure*}
\begin{figure*}
\begin{center}
  \includegraphics[width=1.\textwidth]{tpcf_sigma.pdf}
\caption{1-$\sigma$(dark-shaded) and two-$\sigma$ ABC posterior predictions
for the two-point correlation function of galaxies. The black lines show the data-points.}
\label{fig:gmf_cov}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=1.\textwidth]{tpcf_res_sigma3.pdf}
\caption{1-$\sigma$(dark-shaded) and two-$\sigma$ ABC posterior model prediction residuals
for the two-point correlation function of galaxies.}
\label{fig:2pcf-model}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=1.\textwidth]{gmf3.png}
\caption{{\bf Top Panel:}1-$\sigma$(dark-shaded) and 2-$\sigma$ ABC posterior model prediction 
for the group multiplicity function of galaxies in the mock catalog.{\bf Bottom Panel:}Same as the top panel, but showing the residuals.}
\label{fig:2pcf-model}
\end{center}
\end{figure*}

\section*{Acknowledgments}

We would like to thank Mike Blanton, Boris Leidstadt, Micheal Schneider, 
Uros Seljak, Ross Fadely, Jeremy Tinker, Alex Malz, Patrick McDonald, Dan 
Foreman-Mackey, for productive and insightful discussions.

MJV was supported by NSF grant AST-1517237. CHH was supported by James 
Arthur Graduate Award Grant. DWH was supported by NSF (grants IIS-1124794 
and AST-1517237), NASA (grant NNX12AI50G), and the Moore-Sloan Data Science 
Environment at NYU.

\todo{We need to acknowledge AstroHackWeek.}

\bibliographystyle{yahapj}
\bibliography{ccppabc}
\end{document}