### 2017-03-08 ccppabc/text/response.md

Dear Editor,

This was a very constructive report and we have made revision to the paper
accordingly. We think the comments have led to improvement in quality of the paper.

ChangHooh Hahn & Mohammadjavad Vakili (on behalf of the authors):

-----
We are grateful to the constructive and helpful comments of the 
anonymous referee on our manuscript. They have led to improvements in the paper.  
In what follows, the referee comments are indented and our responses are not indented:
-----

> a deeper discussion on why the ABC approach should be preferred in this particular context 
(one or two phrases on this should be included in the abstract).

>in section 3.4 it is said: "Furthermore, the Gaussian-likelihood approach relies on constructing an accurate covariance
matrix estimate that captures the sample variance of the data. While we are able to
do this accurately within the scope of the HOD framework, for more general LSS parameter
inference situations, it is both labor and computationally expensive and dependent on the
accuracy of simulated mock catalogs, which are known to be unreliable on small scales (see
Heitmann et al. 2008; Chuang et al. 2015 and references therein). Since ABC-PMC utilizes
a forward model to account for sample variance, it does not depend on a covariance matrix
estimate; hence it does not face these problems."
this might lead to confusion. Do the authors believe that ABC presents no advantage  
in the specific case studied in this paper?

> if the Gaussian hypothesis underlying the standard approach are so unrealistic 
as stated in the text, why results are still consistent with the ABC ones? 
how likely it is that we will face a real data scenario which might lead to significantly different results?

> is it possible to simulate an extreme data situation where the results from 
the standard analysis are not consistent with the ABC ones? if so,  how realistic it is?

> what caveats one would face in trying to  apply this kind of analysis in real data? 
How can they affect the final results and how can they be circumvented?